---
title: "dsci301_lecture19"
author: "Dr. Wojciech Golik"
date: "Nov 09, 2022"
output: html_document
---


#Interaction

An interaction occurs when an independent variable has a different effect on the outcome depending on the values of another independent variable.  For example, one variable, x1 might have a different effect on y within different categories or groups given by variable x2.

Syntax:

With lm, an asterisk in the formula generates the interaction terms.

```{r}
#use SEARCH to learn more about state.x77
df <- data.frame(state.x77)

df
```
Compare 
```{r}

mod1 <- lm(Income ~ HS.Grad + Illiteracy + Murder, data = df)

summary(mod1)

```


Predict: HS.Grad = 13,  Illiteracy = 1.5, Murder = 14; can you verify this by hand calculation?

we got the following equation:
Income = 2017.96 + 45.19*13 - 176.84*1.5 + 30.48*14 



```{r}
#evaluating INCOME using PREDICT()
predict(mod1, data.frame(HS.Grad = 13, Illiteracy = 1.5, Murder = 14 ))

#using the formula explicitly:
2017.96 + 45.19*13 - 176.84*1.5 + 30.48*14
```



```{r}
#Fit Income using HS.Grad, Illiteracy, Murder, and interaction of Illiteracy & Murder

mod2 <- lm(Income ~ HS.Grad + Illiteracy*Murder, data = df)

summary(mod2)
```


1. Interpret result

we got the following equation:
Income = 1414.46 + 40.76*13 + 753.07*1.5 + 130.60*14 - 97.04*1.5*14
Note that the interaction between Illiteracy*Murder resulted in the last term!!

2. Predict: HS.Grad = 13, Illiteracy = 1.5, Murder = 14,  can you verify this by hand calculation?
Of course - just plug in the numbers in the formulas above!!

```{r}
predict(mod2, data.frame(  HS.Grad = 13,  Illiteracy = 1.5, Murder = 14))

#using the formula directly...
1414.46 + 40.76*13 + 753.07*1.5 + 130.60*14 - 97.04*1.5*14
```


#Limitation Of Linear Regression

1. the LM models should not be used for EXTRAPOLATION (i.e., outside the data range). 

2. Don't use the LM when target variable is binary. The restriction to 0 and 1 response does not fit the normal distribution.

3. Linear models are sensitive to outliers.

4. They don't perform well with nonlinear relationships, though there may be a data transformation that can produce linearity.


#Now,  download  lab9  from the folder  LABS  and complete it in class... 


