---
title: "dsci301_lecture16"
author: "Dr. Wojciech Golik"
date: "October 24, 2022"
output: html_document
---

Working with linear regression...


```{r}
install.packages("ISLR")
library(ISLR)
library(MASS)

#to read more about the package MASS type MASS  in Packages in RStudio
```


```{r}
#Check out the Boston data set
#to read more about the data 
head(Boston)
dim(Boston)
library("tidyverse")
sample_n(Boston, 10)


```

to read more about lm()  type  "lm"  in  the SEARCH help

```{r}
# Fit a linear regression model
#the variables are:   x  is medv, and  y  is  lstat
mod1 <- lm(medv ~ lstat, data = Boston)
mod1

```


```{r}
#Check out the summary of your model
summary(mod1)
```

```{r}
#plot the points  medv  vs  lstat  with the linear regression 
library(tidyverse)
ggplot(mod1) + 
  geom_point(mapping = aes(x = lstat, y=medv)) + 
  geom_smooth(aes(x = lstat, y=medv),  method = "lm")


```


```{r}
#Use the coef() to determine the coefficients
coef(mod1)
```


```{r}
#Use the confint() to determine the CONFIDENCE INTERVALS of the parameters
confint(mod1, level=0.95)
```

```{r}


plot(c(0, 40), c(0,50) )
abline( a= 34.5538409,   b=-0.9500494) 

abline( a=33.448457 ,   b=-1.026148)
abline( a= 35.6592247,   b=-0.8739505 )


```
```{r}
#we'll use geom_abline() to draw the lines form the confidence intervals...
ggplot(mod1) + 
  geom_point(mapping = aes(x = lstat, y=medv)) + 
  geom_smooth(aes(x = lstat, y=medv),  method = "lm") +
  geom_abline( aes( intercept = 33.448457, slope=-1.026148 ) , alpha=0.5 ) +
  geom_abline( aes( intercept = 35.6592247, slope=-0.8739505 ) , alpha=0.5 )


```

```{r}
#For lstat= c(5 , 10 ,15)), Use the predict() to determine predictions

predict(mod1, data.frame(lstat= c(5 , 10 ,15)))

```

#By hand
```{r}
34.5538409 + -0.9500494 * 5
34.5538409 + -0.9500494 * 10
34.5538409 + -0.9500494 * 15
```
Find the observed values on the LM graph produced above...


```{r}
predict(mod1, data.frame(lstat= c(5 , 10 ,15)), interval='confidence')
```
_____________________________________________________
Lab 9.   Download the file from the FILES/LABS.

1.  Perform the LM analysis (as above ) on medv vs crim.

Write down your insights of the relationship between medv and crim in terms understood by general public.


2.  Perform the LM analysis (as above ) on indus vs crim.
Write down your insights of the relationship between indus and crim in terms understood by general public.







