---
title: "dsci301_lecture15"
author: "Wojciech Golik"
date:     "October 19, 2022"
output: html_document
---

#Exploratory Data Analysis (EDA)
DATAFRAME:    DIAMONDS

```{r}
library(tidyverse)
```

```{r}
dim(diamonds)
head(diamonds)

```




What variable in the diamonds dataset is most important for predicting the price of a diamond? 
we will study the following:
Price vs Carat
Price vs Color
Price vs Clarity
Price vs Cut



1.   First:   Price vs Carat
```{r}
ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_point() + geom_smooth()
```
Generally, the heavier the diamond the more expensive it is.  
However, there are many exceptions to the general trend.   
For example, the are diamonds with carat=1, 1.5, and 3
with prices varying immensely in each weight category...
Other factors influence the price beyond the weight...



2.  Now,   Price vs Color
```{r}
ggplot(data = diamonds, aes(x = color, y = price)) +
  geom_boxplot()
```
The COLOR is coded by letters D-J.  
We note here that the Price median increases with the value of the COLOR letter.   
So do the 1st and 3rd quartiles...
That makes sense...  



3.   Now,  Price vs Clarity

```{r}
ggplot(data = diamonds, aes(x =  clarity, y = price) ) +
  geom_boxplot()
```
On the face of it this is a strange correlation.   
The clearer the diamond the less valuable it becomes...  
Look at the medians...  
How do you explain that??


4. How is price correlated with cut? 
Now    PRICE  vs  CUT
```{r}
ggplot(data = diamonds, aes(x = cut, y = price)) +
  geom_boxplot()
```
Why is the IDEAL cut diamonds less valuable than the ones of any other cut????




5.   What is the relationship between carat and cut? 

```{r}
ggplot(data = diamonds, aes(x = cut, y = carat)) +
  geom_boxplot()
```
What does this relationship between CARAT and CUT  tells us 
about diamonds of lower quality cut being more expensive?

The FAIR cut diamonds are generally larger than the IDEAL cut diamonds.  
That's why (on average) the FAIR diamonds costs more than IDEAL diamonds...
The size (or weight) of diamond is more important that the quality of the cut.

_____________________________________________________________________________________
Exercise.
6.  So what if we compare diamonds in this dataframe of similar weight?
Would the IDEAL diamonds be generally more expensive than the FAIR ones? 
Filter the dataframe DIAMONDS to keep only diamonds with, say, 1.0<= CARAT <=1.5. 
Then do the boxplot of PRICE vs CUT.  What do you see?
Repeat the experiment for other ranges of variable CARAT.



```{r}

```









