---
title: "dsci301_lecture17"
author: "Dr. Wojciech Golik"
---




#Primer on SIMPLE PLOTTING...
```{r}

sale5 <- c(6, 4, 9, 7, 1, 12, 8, 10, 9, 13)
plot(1:10, sale5,    xlim=c(0, 12) ,  ylim=c(0, 15))

abline(a=5, b=2, col ="red")   #plotting line y=5+2x
abline (h=10)
abline( v=8)


```


````{r}
sale5 <- c(6, 4, 9, 7, 1, 12, 8, 10, 9, 13)
plot(1:10, sale5,    xlim=c(0, 12) ,  ylim=c(0, 15))
lsfit(1:10,sale5)
abline(lsfit(1:10,sale5))
abline(lsfit(1:10,sale5, intercept = FALSE), col= "green") # setting intercept=0
#and changing the color of line,  color="green"

```


#SINGLE variable REGRESSION


```{r}
#let's use dataframes from ISLR and MASS
library(ISLR)
library(MASS)
```


```{r}
# Fit a simple linear regression model to data from BOSTON dataframe
head(Boston)

x=   Boston$lstat
y=   Boston$medv
plot(x, y,    xlim=c(0, 40) ,  ylim=c(0, 50))
abline(lsfit(x, y))


#linear regression on data  frames (from TIDYVERSE)
mod1 <- lm(medv ~ lstat, data = Boston)
mod1


```

```{r}
#plot the points  medv  vs  lstat  with the linear regression (using GGPLOT)
library(tidyverse)  #needed  for GGPLOT
ggplot(mod1) + 
  geom_point(mapping = aes(x = lstat, y=medv)) + 
  geom_smooth(aes(x = lstat, y=medv),  method = "lm")

#predict  MEDV   when  lstat=30  using MOD1
predict(mod1, data.frame( lstat=30 )   )



```





```{r}
#plot the points  medv  vs  lstat  with polynomial regression (removing method = "lm") 
library(tidyverse)
ggplot(mod1) + 
  geom_point(mapping = aes(x = lstat, y=medv)) + 
  geom_smooth(aes(x = lstat, y=medv))

```

#Multiple Linear Regression

```{r}
mod1 <- lm(medv ~ lstat + age, data = Boston)
mod1

```

```{r}
#Get summary of mod1
summary(mod1)

```


```{r}
names(Boston)
```

```{r}
#What is ~. doing?

mod2 <- lm(medv ~., data = Boston)
mod2
summary(mod2)
```




Note that   lm(medv ~., data = Boston)  produced the multilinear formula
y = y_intercept + c1*x1 + c2#x2 + c3*x3 + ...
Now - study the p-values of mod2 produced by summary(mod2)  above...

1.  When p-values for individual variables are > than ALPHA (the significance level),
with ALPHA=ranging from 1% to 10%  then that particular variable has (statisticallY) 
NO IMPLAT on the value of the dependent variable and should be REMOVED from further analysis.

2.  Also important is the value of the parameter R-squared  (see above).  Better yet - examine the value of the Adjusted-R-squared coefficients.  It tells us how much of the variation of y is due to variation of the variables we used in our analysis..

3.  But always start with examining the p-Value for the global F-test.   
This value must be very small for the analysis to proceed...

```{r}
#Based on p-values, let's remove predictors that turned out to be not significant from the model.    
mod3 <- lm(medv ~.-indus - age, data = Boston)
summary(mod3)
mod3

```

```{r}
# We can also use the update function, compare both outputs (one above, one below)..
update(mod2, ~. -indus - age)

```


#Model Diagnostics

```{r}
#Create a data frame of predictions and residuals.
#recall that mod3 contains all variables except indus and age
#Plot a scatterplot.




diagnostics <- data.frame(predictions = mod3$fitted.values, residuals = mod3$residuals)


diagnostics %>% 
  ggplot(aes(x = predictions, y = residuals)) +
  geom_point()

#Or

  ggplot(data = diagnostics, mapping = aes(x = predictions, y = residuals)) +
  geom_point()

```




#YOUR TURN

```{r}
#Load the movies data set from canvas (download it from DATA).

movies <-read.csv(file="movies.csv")
movies


```

```{r}
#Fit a linear model: imdb_rating as your target variable and critics_score as predictor.


```



```{r}
#plot the scatter plot and the regression line


```


```{r}
#Add audience_score as additional predictor and fit a lm.


```


```{r}
#Predict the IMDb rating of a model with a critics score of 70 and audience score of 20. 


# predict the IMDb_rating for ALL FITTED VALUES


```


```{r}
#Model diagnostics: Plot predictions  versus residuals
 


```






FINALLY, let's do the model using all the variables
```{r}
movies <- read.csv("movies.csv")
names(movies)
#???
```


```{r}
#Fit linear model imdb_rating using the rest of the predictors.

md <- lm(imdb_rating~. , data = movies)
summary(md)

#???
```









