---
title: "dsci301_lecture13"
author: "Dr. Wojciech Golik"
date: "October 05,2022"
output: html_document
---


Objectives:

At the end of the lesson, students will be able to:

 * use pipe operator  %>%    to summarize datasets.
 * Use group_by to summarize data.
 

```{r}
library(tidyverse)
```

#Pipe Operator %>%

The pipe operator will allow you to avoid either a long nested operation or doing a bunch of assignments.
Numeric example.
```{r}
#Perform the operation below

sin(log(sqrt(sum(c(1, 2, 3, 4) ^ 2))))

```

```{r}
c(1, 2, 3, 4) ^ 2 %>% 
  sum %>% 
  sqrt %>% 
  log %>% 
  sin
```

x %>% f(y) turns into: f(x, y)

x %>% f(y) %>% g(z) turns into: g(f(x, y), z)


Example
```{r}
df <- mtcars

```

```{r}
#Filter mpg > 20
#take a random sample of five observations. 

arrange(sample_n(filter(df,mpg > 20),size=5),desc(mpg))

```


```{r}
df %>% 
  filter(mpg > 20) %>% 
  sample_n(size = 5) %>% 
  arrange(desc(mpg))

```



We have already discussed in LECTURE06  how to:

1. Pick observations by their values: filter()

2. Reorder the rows: arrange()

3. Pick variables by their names: select()

4. Create new variables with functions of existing variables: mutate()

5. Collapse many values down to a single summary: summarise()


All verbs work similarly:

1. The first argument is a tibble (or data frame)

2. The subsequent ones describe what to do, using the variable names

3. The result is a new tibble


```{r}
library(nycflights13)
flights

```

```{r}
sample_n(flights, 10)
```

Example

```{r}
summarise(flights,
          mean_distance = mean(distance, na.rm = TRUE),
          mean_arr_delay = mean(arr_delay, na.rm = TRUE) )
```


# Group by

group_by() takes an existing tbl and converts it into a grouped tbl where operations are performed "by group". ungroup() removes grouping.

grouping doesn't change how the data looks.

It changes how it acts with the other dplyr verbs (summarize, filter, etc.)

```{r}
mtcars
by_cyl <- mtcars %>% group_by(cyl)
by_cyl
summarize(by_cyl,  mean_disp = mean(disp, na.rm = TRUE) )
```


More examples

```{r}
#For the flight data, group by year, month and day
# then find the mean distance and mean arr_delay.
flights
flights %>% 
  group_by(month, day) %>% 
  summarise(mean_dist = mean(distance, na.rm = TRUE),
            mean_arr_delay = mean(arr_delay, na.rm = TRUE)) 


#what is going on here?  explain!
flights %>% 
  group_by(month) %>% 
  summarise(mean_dist = mean(distance, na.rm = TRUE),
            mean_arr_delay = mean(arr_delay, na.rm = TRUE)) 

```





```{r}
#For the flight data, group by destination
#then find Count the number of occurrences using n(), mean distance and mean arr_delay.
#filter those with count > 200 and destination is not HNL.

flights %>% 
  group_by(dest) %>% 
  summarise(count = n(),
            mean_dist = mean(distance, na.rm = TRUE),
            mean_arr_delay = mean(arr_delay, na.rm = TRUE)) %>% 
  filter(count > 200, dest != "HNL")
  

  

```


```{r}
#For the above example, plot a scatterplot mean_dist by mean_arr_delay.
# Add a smoothline and turn off the se.
# What can be inferred from the plot?

flights %>% 
  group_by(dest) %>% 
  summarise(count = n(),
            mean_dist = mean(distance, na.rm = TRUE),
            mean_arr_delay = mean(arr_delay, na.rm = TRUE)) %>% 
  filter(count > 200, dest != "HNL") %>% 
  ggplot(mapping = aes(x = mean_dist, y = mean_arr_delay)) +
  geom_point() +
  geom_smooth(se = FALSE)


```


Exercise:

1.  Use the pipe operator to do the following.  Take the dataframe flights.
Filter it with the distance>1500 miles.
Group it by origin.  Count the number of longer than 1500 miles flights  from each airport.

```{r}




```

2.  Use the pipe operator to do the following.  Take the dataframe flights.
Filter it with the arr_delay>30 minutes.
Group it by carrier.  Count the number of flights with arr_delay>30 minutes by carrier.  





