---
title: "DSCI 301 - Lab 2"
author: "Dillon Geary"
date: "2022-08-29"
output: word_document
---

Now:  do Lab2.
* Create a new R Markdown document with File > New File > R Markdown.
* title: "DSCI 301 - Lab 2"
* author: "Put your name"
* Save your file as  DSCI 301 - Lab 2 - LastName.rmd
* Knit your RMD file into Microsoft Word
* After you are done, load both your files into Canvas, Lab 2, before 12:50p.m



1. Create 2 vectors A and B, where A is (1,2,3) and B is (4,5,6). With these vectors, use the cbind() or rbind() function to create a 2 by 3 matrix C from the vectors. Display C. You'll need to figure out which of these binding functions is the correct choice.

```{r}
A <- c(1:3)
B <- c(4:6)
C<- rbind(A,B)
C
```

2. Create a 3 by 3 matrix consisting of the numbers 1-9. Create this matrix using the shortcut 1:9 and by specifying the nrow argument in the matrix() function call. Assign this matrix to the variable D. Display D.

```{r}
D <- matrix(1:9, nrow=3)
D
```

3. Create a 5 by 5 matrix consisting of the numbers 1-25 and assign it to the variable E. Display E. The top row should be the numbers 1-5.

```{r}
E <- matrix(1:25, nrow=5, byrow= TRUE)
E
```
4. Using indexing notation, grab a sub-section of E from the previous exercise that looks like this:
[7,8]
[12,13]
Assign this subsection to F. Display F.

```{r}
E[2:3,2:3]
```