---
title: "DSCI 301 - Lab 5"
author: "Dillon Geary"
date: "2022-09-20"
output: word_document
---

1. Use the head() function to get to top few rows of the mpg dataset.
Recall that you have to load   library(tidyverse) to have access to mpg.  

```{r}
library(tidyverse)
head(mpg)

```
2. For the mpg data set, plot a scatterplot for displ by hwy.

```{r}
c <- ggplot( data=mpg, mapping = aes( x=displ, y=hwy))

c+ geom_point()

```

3. Facet by class with two rows.
```{r}
c+ geom_point()+ facet_wrap(vars(class),nrow =2)
```
    
4. Facet by class and cyl
```{r}
c+ geom_point()+ facet_wrap(vars(class,cyl),nrow =2)
```


5. From your plot in #4, give any three observations.

* Cars with lower displacement generally get better mpg on a highway
* SUV and Pickups with 8  cylinders tend to have the highest displacement and concurrently the lowest mpg on a highway
* Most cars with similiar displacement less than 3 get the same mpg on a highway on average.
